### 2020-06-25 ArticleAcortadobackup.tex
Final

\usepackage[left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm]{geometry}
\usepackage{graphicx}
\usepackage{amssymb,mathtools, amsmath, amsfonts, amsthm}
%\usepackage{color}
\usepackage{float}
\usepackage{hyperref}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{chngcntr}
\usepackage{cleveref}
\usepackage{pdfpages}

@@

\newlength{\margen}
\setlength{\margen}{\paperwidth}
\addtolength{\margen}{-\textwidth}
\addtolength{\skip\footins}{0.7 cm}
\setlength{\margen}{0.5\margen}
\addtolength{\margen}{-1in}
\setlength{\oddsidemargin}{\margen}
\setlength{\evensidemargin}{\margen}
\setlength{\abovedisplayskip}{3pt}
\setlength{\belowdisplayskip}{3pt}
%%%% Small setup %%%%
\hypersetup{
	colorlinks=false,
	pdfborder={1 1 0.0005},
}
\setlength{\parskip}{0.2cm}
%%%%%%%%%%%%%%%
\usepackage{tikz-cd}
\usetikzlibrary{cd}
\usepackage[english]{babel}
\usepackage{todonotes}
\usepackage{cleveref}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{bbding}
\usepackage{tcolorbox}
\usepackage[square,numbers]{natbib}

@@

\newenvironment{oneshot}[1]{\@begintheorem{#1}{\unskip}}{\@endtheorem}
\newcommand{\cc}{\mathfrak{c}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Hc}{\mathcal{H}}
\newcommand{\Lan}{\mathcal{L}}
\newcommand{\Ln}{\lim\limits_{n\to \infty}}
\newcommand{\clist}{\mathfrak{c}_{1}, \cdots, \mathfrak{c}_m}
\newcommand{\morph}[1]{\sim_#1}
\newcommand{\vlst}[2]{#1_1,\dots, #1_{#2}}
\newcommand{\gnp}{G(n,\beta_1/n^{a_1-1}, \dots,\beta_l/n^{a_l-1})}
\newcommand{\overbar}[1]{\mkern 1.5mu\overline{\mkern-1.5mu#1\mkern-1.5mu}\mkern 1.5mu}
\newcommand{\ehr}{\textsc{Ehr}}
\newcommand{\PR}[1]{\mathrm{Pr}\left(#1\right)}
\newcommand{\Tr}{\mathrm{Tr}}
\newcommand{\InR}[1]{\left\{ #1_R \right\}_{R\in \sigma}}
\newcommand{\sep}{\noindent\rule{2cm}{0.4pt}}
\newcommand{\aut}{\mathrm{aut}}

\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}

@@

We say that a sequence of random structures $\{G_n\}_n$ satisfies a limit
law with respect to some logical language $L$ if for any property $P$ expressible in
$L$ the probability that $G_n$ satisfies $P$ tends to some limit as $n$ goes to $\infty$.
If that limit takes only the values zero and one then we say that $\{G_n\}_n$ satisfies
a zero-one law with respect to $L$.  \par
Convergence and zero one laws have been extensively studied with regards to 
random binomial graphs $G(n,p)$.
The seminal theorem on this topic, due to Fagin \cite{fagin1976probabilities}
and Glebskii et al. \cite{glebskii1969range} independently, 
concerns general relational structures.
When applied to graphs it states that if $p$ is a fixed probability
then $G(n,p)$ satisfies a zero-one law with respect to the first order
(FO) language of graphs. \par
This zero-one law was later extended by Shelah and Spencer
in \cite{shelah1988zero}. There it is proven, among other results,
that if $p:=p(n)$ was a decreasing function on
$n$ of the form $n^{-\alpha}$, and $\alpha>0$ is irrational
then $G(n,p(n))$ obeys a zero-one law with respect to FO logic. Moreover, 
it is also proven that if $\alpha\in (0,1)$ is rational then $G(n,p(n))$ 
does not obey a convergence law. \par
This was further extended by Lynch
in \cite{lynch1992probabilities}, where it is shown that in the
case where the expected number of edges is linear, when $p(n)\sim \beta/n$
for some $\beta\in (0,\infty)$, then $G(n,p(n))$ satisfies a limit law with
respect to FO logic. The following is a restatement of the main theorem
in that article.

\begin{theorem*}[Lynch, 1992]
	Let $p(n)\sim \beta/n$. For any FO sentence $\phi$, the function
	$F_\phi: (0,\infty)\rightarrow [0,1]$ given by 
	\[ F_\phi(\beta) = \Ln \mathrm{Pr}\left( \mathrm{G}(n,p(n))
	\text{ satisfies } \phi   \right) \]
	is well defined and is given by an expression with parameter $\beta$ built
	using rational constants, addition, multiplication and exponentiation with base $e$.
\end{theorem*}

A relevant aspect of this result is that the limit probability of any FO
property in $G(n,p(n))$ when $p(n)\sim \beta/n$ varies analytically with $\beta$.
A consequence of this is that in these conditions no FO property can "capture" sudden changes 
in the structure of $G(n,p(n))$. \par
It was left open at the end of \cite{lynch1992probabilities} whether the convergence law obeyed
by $G(n,p(n))$in the range $p(n)\sim \beta/n$
could be generalized for other random models of relational structures 
that contain relations of arity greater than $2$.
A result in this regard was obtained in \cite{salvadorbrasil}, 
among other zero-one and convergence laws.
There $G^d(n,p)$ is considered, the random model of $d$-uniform hypergraphs where
each $d$-edge is added to a set of $n$ labeled vertices with probability $p$ independently.
It is shown that when $p(n)\sim \beta/n^{d-1}$ -i.e., when the expected number of
edges is linear- then $G^d(n,p(n))$ obeys a convergence law with respect of the FO
language of $d$-uniform hypergraphs. With little additional work it can be shown that
in these conditions the limit probability of any FO property of $G^d(n,p(n))$ varies 
analytically with $\beta$. We 
extend this result to arbitrary relational structures
to whose relations we can impose symmetry and anti-reflexivity constraints 
(\cref{thm:main}). \par

This generalization is motivated by an application to the problem of random 
satisfiability. We continue the study started by Atserias in 
\cite{atserias2005definability}
with regards to the definability in first order
logic of certificates for unsatisfiability
that hold for typical unsatisfiable formulas. 
A random model for $3$-CNF formulas where 
each possible clause over $n$ variables 
is added with probability $p$ independently is considered there.
This way the expected number of clauses $m$ is $\Theta(n^3p)$ as $n$ grows.
The main result of that article states the following:
(1) if $m=\Theta(n^{2-\alpha})$ 
for any irrational $\alpha>0$, then no FO property of $3$-CNF formulas
that implies unsatisfiability holds asymptotically almost surely (a.a.s) 
for unsatisfiable formulas, and (2) if $m=\Theta(n^{2+\alpha})$ for any $\alpha>0$,
then there exists some FO property that implies
unsatisfiability and  holds a.a.s for unsatisfiable formulas. \par
The second part of the statement is the simpler one to prove:
it can be shown that when $m=\Theta(n^{2+\alpha})$ for some 
$\alpha>0$ then the random $3$-CNF formula
a.a.s contains some fixed unsatisfiable subformula (which depends
on the choice of $\alpha$). This is clearly expressible in FO logic, 
so (2) follows. The proof of (1) is more involved and, in fact, 
shows something stronger: if $m=\Theta(n^{2-\alpha})$
for any $\alpha>0$ irrational, then all FO properties 
that imply unsatisfiability
a.a.s do not hold. This proof employs
techniques based in the ones used by Shelah and Spencer 
in \cite{shelah1988zero} to prove that $G(n,p)$ satisfies a 
zero-one law with respect to FO logic when $p$ is an irrational 
power of $n$. \par
Since the techniques used to prove (2) rely in the fact that $\alpha$
is irrational, the study of the range $m=\Theta(n)$
(that is, $m=\Theta(n^{2-\alpha})$ with $\alpha=1$), was left open.
This range is of special interest because it is where the phase 
transition from almost sure satisfiability to almost sure unsatisfiability
takes place. It was shown
in \cite{chvatal1992mick} that a random $k$-CNF formula with
$m$ clauses over $n$ variables satisfying that $m\sim c\cdot n$ 
is a.a.s satisfiable for all sufficiently small values of $c$ and is a.a.s
unsatisfiable for all sufficiently big values of $c$.
\par
The possibility of studying FO definability of certificates for unsatisfiability
in random $l$-CNF formulas with a linear expected number of clauses 
using a generalization of Lynch's theorem was suggested by Atserias. 
This application is discussed in \cref{sec:SAT}. We give a brief overview
of it here. Let $F(l,n,p)$ be a random model of $l$-CNF 
formulas where each $l$-clause over $n$ variables is chosen with probability
$p$ independently. Let $F_n(\beta)$ denote a random formula in
$F(l,n,p)$ where $p:=p(n)\sim \beta/n^{l-1}$. 
Suppose that 
every FO property of $l$-CNF formulas 
has a well defined asymptotic probability in $F_n(\beta)$
for any $\beta>0$.
Further suppose that those asymptotic probabilities vary analytically
with $\beta$. Then any FO property that implies unsatisfiability
a.a.s does not hold in $F_n(\beta)$ for any
$\beta>0$. Indeed, let $P$ be one of such FO properties. One can find a
value $\beta_0>0$ satisfying that 
a.a.s $F_n(\beta)$ is satisfiable when $0<\beta<\beta_0$.
In consequence $P$ a.a.s does not hold
in $F_n(\beta)$ when $0<\beta<\beta_0$.
Since the asymptotic probability of $P$ varies analytically
with $\beta$ and it vanishes in the non-empty interval $(0,\beta_0)$,
because of the Principle of Analytical Continuation
it must be true that a.a.s $P$ does not hold in $F_n(\beta)$ for all
$\beta>0$.

@@

Given a positive natural number $n$, we will write
$[n]$ to denote the set ${1,2,\dots,n}$.\par
Given numbers, $n,m\in \N$ with $m\leq n$ we denote by
$(n)_m$ the $m$-th falling factorial of $n$. \par
Given a set $S$ and a natural number $k\in \N$
we will use $\binom{S}{k}$ to denote the set of 
subsets of $S$ whose size is $k$.Given a set
$S$ and a number $n\in \N$ with $n\leq |S|$ we define
$(S)_n$ as the subset of $S^n$ consisting of the $n$-tuples
whose coordinates are all different. 
We also define $S^*:=\bigcup_{n=0}^\infty S^n$ and
$(S)_*:=\bigcup_{n\leq |S|} (S)_n$. \par

We will use the convention that over-lined variables
, like $\overline{x}$, denote ordered tuples of arbitrary length.
Given an ordered tuple $\overline{x}$
we define the number $len(\overline{x})$ as its length. 
Given a tuple $\overline{x}$ and an element $x$ the expression
$x\in \overline{x}$ means that $x$ appears as some coordinate
in $\overline{x}$. 
Given a map $f:X\rightarrow Y$ between two sets $X, Y$ and 
an ordered tuple $\overline{x}:=(x_1,\dots,x_a)\in X^*$ 
we define $f(\overline{x})\in Y^*$ as the tuple 
$(f(x_1),\dots,f(x_a))$.
Given two tuples $\overline{x},\overline{y}$
we write $\overline{x}^\smallfrown \overline{y}$ to denote their 
concatenation. Given a set $S$ and elements $x_s$ for each $s\in S$
we will write $\{x_s\}_{s\in S}$, or just $\{x_s\}_s$ when $S$ is understood,
to denote the tuple indexed by $S$ which for any $s\in S$
contains the element $x_s$ at the position given by $s$. 
\par

Let $S$ be a set, $a$ a positive natural number, 
and $\Phi$ a group of permutations over 
$[a]$. Then $\Phi$ acts naturally over
$S^a$ in the following way: Given $g\in \Phi$ and
$\overline{x}:=(x_1,\dots,x_a)\in S^a$ we define 
$g\cdot \overline{x}$ 
as the tuple $(x_{g(1)},\dots,x_{g(a)})$. 
We will denote by $S^a/\Phi$ to the quotient
of the set $S^a$ by this action. Given an element
$\overline{x}:=(x_1,\dots, x_a)\in S^a$ we denote its equivalence
class in $S^a/\Phi$ by $[x_1,\dots,x_a]$ or $[\overline{x}]$.
Thus, for any
$g\in \Phi$, by definition $[x_1,\dots,x_a]=[x_{g(1)}
,\dots,x_{g(a)}]$. \par
The notations $\overline{x}$ and
$(x_1,\dots, x_a)$ 
represent
ordered tuples while 
$[\overline{x}]$ and
$[x_1,\dots,x_a]$ denote ordered tuples modulo the
action of some arbitrary group of permutations. Which group is
this will depend on the ambient set where $[x_1,\dots,x_a]$ belongs
and it should either be clear from context or not be relevant.\par

Given real functions over the natural numbers 
$f,g:\N \rightarrow \R$ the expressions $f=O(g)$,
$f=o(g)$ and $f=\Theta(g)$ will have their usual meaning.
If $g(n)\neq 0$ for $n$ large enough then
we will write $f\sim g$ when $\Ln \frac{f(n)}{g(n)}=1$. \par

@@

We assume familiarity with basic probability theory. 
We denote by $\mathrm{Poiss}_\lambda(n)$ the discrete probability mass
function of a random variable following a Poisson distribution with
mean $\lambda$. That is, $\mathrm{Poiss}_\lambda(n)=e^{-\lambda}\frac{\lambda^n}{n!}$.
We define $\mathrm{Poiss}_\lambda(\geq n)=1 - \sum_{i=0}^{n-1} \mathrm{Poiss}_\lambda(i)$. \par

@@

We will use the following observation in order to compute the binomial moments of 
our random variables.

\begin{obs} \label{obs:binomialmean} Let $X_1,\dots, X_l$ be non negative
	random integer variables over the same space. 
	Let $r_1,\dots,r_l\in \N$.	Suppose
	each $X_i$ is the sum of various indicator random variables
	(i.e. variables that only take the values $0$ and $1$)
	$X_i=\sum_{j=1}^{a_i} Y_{i,j}$. Define 
	$\Omega:=\prod_{i=1}^l \binom{[a_i]}{r_i}$. That is,
	the elements $\{S_i\}_{i\in[l]}\in \Omega$
	represent all the possible unordered choices of 
	$r_i$ indicator variables $Y_{i,j}$ for each $i\in [l]$.
	Then 
	\[
	\mathrm{E}\left[
	\prod_{i=1}^{l} \binom{X_i}{r_i}\right]=
	\sum_{\{S_i\}_{i\in [l]}} \mathrm{Pr}\left(
	\bigwedge_{\substack{i\in [l]\\ j\in S_i}} Y_{i,j}=1
	\right).	
	\]
\end{obs}

If $A,B$ are events we may write the conditioned probability $\mathrm{Pr}(A\, | \, B )$ 
as $\mathrm{Pr}_B(A)$ in some situations to shorten some expressions.

\subsection{Logical preliminaries}
We assume familiarity with first order logic (FO). We follow 
the convention that first order logic contains the equality symbol. 
Given a vocabulary $\sigma$ we will denote by $FO[\sigma]$ the set of 
first order formulas of vocabulary $\sigma$. We define the set of 
\textbf{free variables} of a formula as usual. 
Given a relation symbol $R\in \sigma$ we denote by $ar(R)$ the arity of $R$. 
Given a formula $\phi\in FO[\sigma]$ we will use the notation $\phi(\overline{y})$ 
to express that $\overline{y}$ is a tuple of 
(different) variables which contains all free variables in $\phi$ and
none of its bounded variables, but it may contain variables
which do not appear in $\phi$.
Formulas with no free variables are called \textbf{sentences} and 
formulas with no quantifiers are called \textbf{open formulas}. 
We call \textbf{edge sentence} to any consistent open formula that contains
no occurrence of the equality symbol `$=$'.

@@

For the rest of the article consider fixed:
\begin{itemize}
	\item A relational vocabulary $\sigma$ such 
	that all the relations $R\in\sigma$ satisfy $ar(R)\geq 2$. 
	\item 
	Groups $\{ \Phi_R \}_{R\in \sigma}$
	such that each $\Phi_R$ is consists of 
	permutations on $[ar(R)]$ with the usual 
	composition as its operation.	
	\item 
	Sets $\{P_R\}_{R\in \sigma}$ satisfying that for 
	all $R\in \sigma$, 
	$P_R\subseteq \binom{[ar(R)]}{2}$

@@

We define the class $\mathcal{C}$ as the 
class of $\sigma$-structures that
satisfy the 
following axioms: 
\begin{itemize}
	\item \textit{Symmetry axioms}: For each $R\in \sigma$ and
	each $g\in \Phi_R$:
	\[ \forall \overline{x}:=x_1,\dots, x_{ar(R)} \left(  R(\overline{x})
	\iff R(g\cdot\overline{x}) \right)    \]
	\item \textit{Anti-reflexivity axioms}: For each 
	$R\in \sigma$ and $\{i,j\}\in P_R$
	\[ \forall x_1,\dots, x_{ar(R)} 
	\left( (x_i=x_j) \implies \neg R(x_1,\dots, x_{a_s})
	\right)\]
\end{itemize}

>>>>

\usepackage[left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm]{geometry}
\usepackage{graphicx}
\usepackage{amssymb,mathtools, amsmath, amsfonts, amsthm}
%\usepackage{color}
\usepackage{float}
\usepackage{hyperref}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{chngcntr}
\usepackage[noabbrev]{cleveref}
\usepackage{pdfpages}

@@

\newlength{\margen}
\setlength{\margen}{\paperwidth}
\addtolength{\margen}{-\textwidth}
\addtolength{\skip\footins}{0.7 cm}
\setlength{\margen}{0.5\margen}
\addtolength{\margen}{-1in}
\setlength{\oddsidemargin}{\margen}
\setlength{\evensidemargin}{\margen}
\setlength{\abovedisplayskip}{2pt}
\setlength{\belowdisplayskip}{2pt}
%%%% Small setup %%%%
\hypersetup{
	colorlinks=false,
	pdfborder={1 1 0.0005},
}
\setlength{\parskip}{0.1 cm}
%%%%%%%%%%%%%%%
\usepackage{tikz-cd}
\usetikzlibrary{cd}
\usepackage[english]{babel}
\usepackage{todonotes}
\usepackage{cleveref}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{bbding}
\usepackage{tcolorbox}
\usepackage[square,numbers]{natbib}

@@

\newenvironment{oneshot}[1]{\@begintheorem{#1}{\unskip}}{\@endtheorem}
\newcommand{\cc}{\mathfrak{c}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Hc}{\mathbf{H}}
\newcommand{\Lan}{\mathcal{L}}
\newcommand{\Ln}{\lim\limits_{n\to \infty}}
\newcommand{\clist}{\mathfrak{c}_{1},  s, \mathfrak{c}_m}
\newcommand{\morph}[1]{\sim_#1}
\newcommand{\vlst}[2]{#1_1,\dots, #1_{#2}}
\newcommand{\gnp}{G(n,\beta_1/n^{a_1-1}, \dots,\beta_l/n^{a_l-1})}
\newcommand{\overbar}[1]{\mkern 1.5mu\overline{\mkern-1.5mu#1\mkern-1.5mu}\mkern 1.5mu}
\newcommand{\ehr}{\textsc{Ehr}}
\newcommand{\PR}[1]{\mathrm{Pr}\left(#1\right)}
\newcommand{\Tr}{\mathrm{Tr}}
\newcommand{\InR}[1]{\left\{ #1_R \right\}_{R\in \sigma}}
\newcommand{\sep}{\noindent\rule{2cm}{0.4pt}}
\newcommand{\aut}{\mathrm{aut}}
\newcommand{\len}{\mathrm{len}}
\newcommand{\qr}{\mathrm{qr}}
\newcommand{\ex}{\mathrm{ex}}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}

@@

We say that a sequence of random structures $\{G_n\}_n$ satisfies a limit
law with respect to some logical language $L$ if for every property $P$ expressible in
$L$ the probability that $G_n$ satisfies $P$ tends to some limit as $n \to \infty$.
If that limit takes only the values zero and one then we say that $\{G_n\}_n$ satisfies
a zero-one law with respect to $L$.  \par
Convergence and zero one laws have been extensively studied on the binomial graph $G(n,p)$.
The seminal theorem on this topic, due to Fagin \cite{fagin1976probabilities}
and Glebskii et al. \cite{glebskii1969range} independently, 
concerns general relational structures.
When applied to graphs it states that if $p$ is fixed,
then $G(n,p)$ satisfies a zero-one law with respect to the first order
(FO) language of graphs. \par
This zero-one law was later extended by Shelah and Spencer
in \cite{shelah1988zero}. There it is proven, among other results,
that if $p:=p(n)$ is a decreasing function on
$n$ of the form $n^{-\alpha}$ and $\alpha>0$ is irrational,
then $G(n,p(n))$ obeys a zero-one law with respect to FO logic. Moreover, 
it is also proven that if $\alpha\in (0,1)$ is rational then $G(n,p(n))$ 
does not obey a convergence law. \par
This was further extended by Lynch
in \cite{lynch1992probabilities}, where it is shown that in the
case where the expected number of edges is linear, i.e. when $p(n)\sim \beta/n$
for some $\beta >0$, then $G(n,p(n))$ satisfies a limit law with
respect to FO logic. The following is a restatement of the main result
in that article.

\begin{theorem*}[Lynch, 1992]
	Let $p(n)\sim \beta/n$. For every FO sentence $\phi$, the function
	$F_\phi: (0,\infty)\rightarrow [0,1]$ given by 
	\[ F_\phi(\beta) = \Ln \mathrm{Pr}\left( \mathrm{G}(n,p(n))
	\text{ satisfies } \phi   \right) \]
	is well defined and is given by an expression with parameter $\beta$ built
	using rational constants, addition, multiplication and exponentiation with base $e$.
\end{theorem*}

A relevant aspect of this result is that the limit probability of any FO
property in $G(n,p(n))$ when $p(n)\sim \beta/n$ varies analytically with $\beta$.
A consequence of this is that FO logic cannot "capture" sudden changes 
in the structure of $G(n,p(n))$. \par
It was left open at the end of \cite{lynch1992probabilities} whether the convergence law obeyed
by $G(n,p(n))$in the range $p(n)\sim \beta/n$
could be generalized for other random models of relational structures 
that contain relations of arity greater than $2$.
A result in this regard was obtained in \cite{salvadorbrasil}, 
among other zero-one and convergence laws.
There is considered the random model of $d$-uniform hypergraphs $G^d(n,p)$ where
each $d$-edge is added to a set of $n$ labeled vertices with probability $p$ independently.
It is shown that when $p(n)\sim \beta/n^{d-1}$, i.e. when the expected number of
edges is linear, then $G^d(n,p(n))$ obeys a convergence law with respect to the FO
language of $d$-uniform hypergraphs. With little additional work it can be shown that
in these conditions the limit probability of any FO property of $G^d(n,p(n))$ varies 
analytically with $\beta$. We 
extend this result to arbitrary relational structures
on whose relations we can impose symmetry and anti-reflexivity constraints 
(\Cref{thm:main}). \par

This generalization is motivated by an application to the problem of random 
SAT. We continue the study started by Atserias in 
\cite{atserias2005definability}
with regards to the definability in first order
logic of certificates for unsatisfiability
that hold for typical unsatisfiable formulas. 
A random model for $3$-CNF formulas where 
each possible clause over $n$ variables 
is added with probability $p$ independently is considered there.
This way the expected number of clauses $m$ is $\Theta(n^3p)$ as $n$ grows.
The main result of that article states the following:
(1) if $m=\Theta(n^{2-\alpha})$ 
for any irrational $\alpha>0$, then no FO property of $3$-CNF formulas
that implies unsatisfiability holds asymptotically almost surely (a.a.s) 
for unsatisfiable formulas, and (2) if $m=\Theta(n^{2+\alpha})$ for any $\alpha>0$,
then there exists some FO property that implies
unsatisfiability and  holds a.a.s for unsatisfiable formulas. \par
The second part of the statement is the simpler one to prove:
it can be shown that when $m=\Theta(n^{2+\alpha})$ for some 
$\alpha>0$ then the random $3$-CNF formula
a.a.s contains some fixed unsatisfiable subformula (which depends
on the choice of $\alpha$). This is clearly expressible in FO logic, 
so (2) follows. The proof of (1) is more involved and, in fact, 
shows something stronger: if $m=\Theta(n^{2-\alpha})$
for any $\alpha>0$ irrational, then all FO properties 
that imply unsatisfiability
a.a.s do not hold. This proof employs
techniques based in the ones used by Shelah and Spencer 
in \cite{shelah1988zero} to prove that $G(n,p)$ satisfies a 
zero-one law with respect to FO logic when $p$ is an irrational 
power of $n$. \par
Since the techniques used to prove (2) rely in the fact that $\alpha$
is irrational, the study of the range $m=\Theta(n)$
(that is, $m=\Theta(n^{2-\alpha})$ with $\alpha=1$), was left open.
This range is of special interest because it is where the phase 
transition from almost sure satisfiability to almost sure unsatisfiability
takes place. It was shown
in \cite{chvatal1992mick} that a random $k$-CNF formula with
$m$ clauses over $n$ variables satisfying that $m\sim c  n$ 
is a.a.s satisfiable for all sufficiently small values of $c$ and is a.a.s
unsatisfiable for all sufficiently big values of $c$.
\par
The possibility of studying FO definability of certificates for unsatisfiability
in random $l$-CNF formulas with a linear expected number of clauses 
using a generalization of Lynch's theorem was suggested by Atserias. 
This application is discussed in \Cref{sec:SAT}. We give a brief overview
of it here. Let $F(l,n,p)$ be a random model of $l$-CNF 
formulas where each $l$-clause over $n$ variables is chosen with probability
$p$ independently. Let $F_n(\beta)$ denote a random formula in
$F(l,n,p)$ where $p:=p(n)\sim \beta/n^{l-1}$. 
Suppose that 
every FO property of $l$-CNF formulas 
has a well defined asymptotic probability in $F_n(\beta)$
for any $\beta>0$.
Further suppose that those asymptotic probabilities vary analytically
with $\beta$. Then any FO property that implies unsatisfiability
a.a.s does not hold in $F_n(\beta)$ for any
$\beta>0$. Indeed, let $P$ be one of such FO properties. One can find a
value $\beta_0>0$ satisfying that 
a.a.s $F_n(\beta)$ is satisfiable when $0<\beta<\beta_0$.
In consequence $P$ a.a.s does not hold
in $F_n(\beta)$ when $0<\beta<\beta_0$.
Since the asymptotic probability of $P$ varies analytically
with $\beta$ and it vanishes in the non-empty interval $(0,\beta_0)$,
because of the Principle of Analytical Continuation
it must be true that a.a.s $P$ does not hold in $F_n(\beta)$ for all
$\beta>0$.

@@

Given a positive natural number $n$, we write
$[n]$ to denote the set ${1,2,\dots,n}$.
Given numbers, $n,m\in \N$ with $m\leq n$ we denote by
$(n)_m$ the $m$-th falling factorial of $n$. 
Given a set $S$ and a natural number $k\in \N$
we use $\binom{S}{k}$ to denote the set of 
subsets of $S$ of size $k$.Given a set
$S$ and $n\leq |S|$, we define
$(S)_n$ as the subset of $S^n$ consisting of the $n$-tuples
whose coordinates are all different. 
We also define $S^*:=\bigcup_{n=0}^\infty S^n$ and
$(S)_*:=\bigcup_{n\leq |S|} (S)_n$. \par

We use the convention that over-lined variables, 
like $\overline{x}$, denote ordered tuples of arbitrary length.
Given an ordered tuple $\overline{x}$
we definer $\len(\overline{x})$ as its length. 
Given a tuple $\overline{x}$ and an element $x$ the expression
$x\in \overline{x}$ means that $x$ appears as some coordinate
in $\overline{x}$. 
Given a map $f:X\rightarrow Y$ and 
an ordered tuple $\overline{x}:=(x_1,\dots,x_a)\in X^*$ 
we define $f(\overline{x})\in Y^*$ as the tuple 
$(f(x_1),\dots,f(x_a))$.
Given two tuples $\overline{x},\overline{y}$
we write $\overline{x}^\smallfrown \overline{y}$ to denote their 
concatenation. Given a set $S$ and elements $x_s$ for each $s\in S$
we write $\{x_s\}_{s\in S}$, or just $\{x_s\}_s$ when $S$ is understood,
to denote the tuple indexed by $S$ which
contains the element $x_s$ at the position given by $s$. 
\par
Let $S$ be a set, $a$ a positive natural number, 
and $\Phi$ a group of permutations over 
$[a]$. Then $\Phi$ acts naturally on
$S^a$ in the following way: Given $g\in \Phi$ and
$\overline{x}:=(x_1,\dots,x_a)\in S^a$ let 
$g  \overline{x}:=(x_{g(1)},\dots,x_{g(a)})$. 
We denote by $S^a/\Phi$ the quotient
of $S^a$ by this action. Given
$\overline{x}:=(x_1,\dots, x_a)\in S^a$ we denote its equivalence
class in $S^a/\Phi$ by $[x_1,\dots,x_a]$ or $[\overline{x}]$.
Thus, for
$g\in \Phi$, by definition $[x_1,\dots,x_a]=[x_{g(1)}
,\dots,x_{g(a)}]$. \par
The notations $\overline{x}$ and
$(x_1,\dots, x_a)$ 
represent
ordered tuples while 
$[\overline{x}]$ and
$[x_1,\dots,x_a]$ denote ordered tuples modulo the
action of some arbitrary group of permutations. Which group it is
will depend on the ambient set where $[x_1,\dots,x_a]$ belongs
and it should either be clear from context or not be relevant.\par

Given real functions over the natural numbers 
$f,g:\N \rightarrow \R$ the expressions $f=O(g)$,
$f=o(g)$ and $f=\Theta(g)$ have their usual meaning.
If $g(n)\neq 0$ for $n$ large enough 
we write $f\sim g$ if $\Ln \frac{f(n)}{g(n)}=1$. \par

@@

We assume familiarity with basic probability theory. 
We denote by $\mathrm{Poiss}_\lambda(n)$ the discrete probability mass
function of a random variable following a Poisson distribution with
mean $\lambda$. That is, $\mathrm{Poiss}_\lambda(n)=e^{-\lambda}\frac{\lambda^n}{n!}$.
We define $\mathrm{Poiss}_\lambda(\geq n)=1 - \sum_{i=0}^{n-1} \mathrm{Poiss}_\lambda(i)$. \par
Given some sequence of events $\{A_n\}_n$ we say that $A_n$ is 
satisfied asymptotically
almost surely (a.a.s.) if $\Pr(A_n)$ tends to $1$ as $n\to \infty$.\par
Given a sequence of random variables $\{X_n\}_n$, the \textbf{first moment method}
is an application of Markov's inequality that establishes that if $\mathrm{E}[X_n]$
tends to zero as $n\to \infty$ then a.a.s $X_n=0$.\par

@@

We use the following observation in order to compute the binomial moments of 
our random variables.

\begin{obs} \label{obs:binomialmean} Let $X_1,\dots, X_l$ be non negative
	random integer variables over the same probability space. 
	Let $r_1,\dots,r_l\in \N$.	Suppose
	each $X_i$ is the sum of indicator random variables
	(i.e. variables that only take the values $0$ and $1$)
	$X_i=\sum_{j=1}^{a_i} Y_{i,j}$. Define 
	$\Omega:=\prod_{i=1}^l \binom{[a_i]}{r_i}$. That is,
	the elements $\{S_i\}_{i\in[l]}\in \Omega$
	represent all the possible unordered choices of 
	$r_i$ indicator variables $Y_{i,j}$ for each $i\in [l]$.
	Then 
	\[
	\mathrm{E}\left[
	\prod_{i=1}^{l} \binom{X_i}{r_i}\right]=
	\sum_{\{S_i\}_{i\in [l]}} \mathrm{Pr}\left(
	\bigwedge_{\substack{i\in [l]\ j\in S_i}} Y_{i,j}=1
	\right).	
	\]
\end{obs}

If $A,B$ are events we may write the conditioned probability $\mathrm{Pr}(A\, | \, B )$ 
as $\mathrm{Pr}_B(A)$ in some situations to shorten some expressions. In this situation, 
given a random variable $X$ we put $\mathrm{E}_B[X]$ to denote conditional expectation of $X$ given the event $B$.

\subsection{Logical preliminaries}
We assume familiarity with first order logic (FO). We follow 
the convention that first order logic contains the equality symbol. 
Given a vocabulary $\sigma$ we denote by $FO[\sigma]$ the set of 
first order formulas of vocabulary $\sigma$.
Given a relation symbol $R\in \sigma$ we denote by $ar(R)$ the arity of $R$. 
Given a formula $\phi\in FO[\sigma]$ we use the notation $\phi(\overline{y})$ 
to express that $\overline{y}$ is a tuple of 
(different) variables which contains all free variables in $\phi$ and
none of its bounded variables, but it may contain variables
which do not appear in $\phi$.
Formulas with no free variables are called \textbf{sentences} and 
formulas with no quantifiers are called \textbf{open formulas}. The \textbf{quantifier 
rank} of a formula $\phi$, written as $\qr(\phi)$, is the maximum number of 
nested quantifiers in $\phi$.
We call \textbf{edge sentence} to any consistent open formula that contains
no occurrence of the equality symbol `$=$'.

@@

For the rest of the article consider fixed:
\begin{itemize}
	\item A relational vocabulary $\sigma$ such 
	that all the relations $R\in\sigma$ satisfy $ar(R)\geq 2$. 
	\item 
	Groups $\{ \Phi_R \}_{R\in \sigma}$
	such that each $\Phi_R$ is consists of 
	permutations on $[ar(R)]$ with the usual 
	composition as its operation.	
	\item 
	Sets $\{P_R\}_{R\in \sigma}$ satisfying
	$P_R\subseteq \binom{[ar(R)]}{2}$ for 
	all $R\in \sigma$.

@@

We define $\mathcal{C}$ as the 
class of $\sigma$-structures that
satisfy the 
following axioms: 
\begin{itemize}
	\item \textit{Symmetry axioms}: For each $R\in \sigma$ and
	$g\in \Phi_R$:
	\[ \forall \overline{x}:=x_1,\dots, x_{ar(R)} \left(  R(\overline{x})
	\iff R(g \overline{x}) \right)    \]
	\item \textit{Anti-reflexivity axioms}: For each 
	$R\in \sigma$ and $\{i,j\}\in P_R$
	\[ \forall x_1,\dots, x_{ar(R)} 
	\left( (x_i=x_j) \implies \neg R(x_1,\dots, x_{a_s})
	\right)\]
\end{itemize}
